\papertitle

### 2014-03-20 final-report/paper.tex
Correct part about colors, remove word span

In particular, \system is built to inspect query execution in
Myria~\cite{myria}, a distributed big data
management system currently being developed in the UW CSE database group. Myria
aims towards building a distributed database platform to provide \emph{big data
management and analytics as a service} primarily for scientific applications.
The proposed visualizations can easily be applied to other DDBMS as well (e.g.
Spark, Hadoop).

@@

Our goal is to build a visualization tool designed to help a database administrator understand the data flow and run-time
performance during query execution in a Distributed Database Management System (DDBMS). Using our tool, the
database programmer or administrator could quickly discover run-time bottlenecks as well as problematic data skew issues,
saving them significant guesswork and programming effort to identify the appropriate logs to analyze. Our visualization tool,
\system, uses profiling logs from Myria as the prototype DDMS. Myria is a distributed
database platform that provides \emph{big data management and analytics as a service} primarily for scientific applications.
Myria takes queries in high-level query languages, such as SQL, translating them into physical query plans and finally executing
them in parallel. Currently, Myria does not offer a visualization tool designed to show the physical query plan, job execution and network
communication between distributed workers. \system allows the programmer/database administrator to visualize the following aspects
of query execution.
\todo{it's not a dag, as we noticed}

@@

Ambrose\cite{ambrose} is a platform developed by Twitter to visualize and
real-time monitor MapReduce data workflows. Ambrose
offers three different views to show associated jobs, job dependencies and
progress, which are not suitable for our needs as the abstraction level of jobs
is too high and does not capture single operators.

Google's Dapper\cite{sigelman2010dapper} is a distributed systems tracing
infrastructure that offers fine-grained tracing of calls in Google's distributed
systems. They also proposed an interface for visualizing traces. Similarly,
X-trace\cite{fonseca2007x} was developed as a framework to trace which events
cause what other events in a distributed environment. Recently, there has been
work on visualizing event traces collected in
X-trace\footnote{\url{https://github.com/brownsys/X-Trace/tree/master/src/webui/html/interactive}}.
Due to the importance of these kinds of debug facilities, Twitter closely
modeled Zipkin\cite{zipkin} after Dapper and X-trace and released it as open
source.

Dapper and X-trace focus on how data flows through a distributed system. In
contrast, in \system the visualization focuses on the operators and how the
data flows through them. This orthogonal view is better suited for debugging
performance bottlenecks in DDBMSs and also scales to a larger number of events.
Furthermore, \system offers different abstraction levels, which enables users
to find problems faster and handle larger amounts of profiling data. Also,
\system is specifically designed to help developers understand query execution
in distributed database system as opposed to general traces in distributed
systems.

Tools to visualize query plans, for example those used to improve the
performance for the SDSS Sky survey\cite{szalay2002sdss}, focus on optimizing
queries and not query execution and have no visualization of data flow, which
is necessary to optimize physical query execution.

@@

\todo{use the word spans when referring to events with duration (used in dapper)}

@@

The Worker Communication view consists of two charts: the aggregate communication
chart and the time series chart. The aggregate communication chart is a matrix representing
a cluster of $N$ workers. Each cell $m,n$ in the matrix represents the total number of tuples sent
by worker $m$ to worker $n$ during the query execution. Figure~\ref{fig:matrix} shows this chart along
with the selected link in the query plan DAG. Hovering the mouse over a cell brings up a tooltip
stating the total number of tuples exchanged between the source and destination workers. The intensity
of the cell color reflects the number of tuples exchanged, i.e., a larger number translates into a darker color
and a smaller number results in a lighter color. The color pallet used is grayscale and hence
immune to color-blindness issues. The color scale uses LAB interpolation.
The matrix chart representing aggregate communication
between workers help uncover data skew issues, as discussed in \S \ref{sec:eval}.

We initially considered having a chord diagram\footnote{\url{http://bl.ocks.org/mbostock/4062006}} instead of a matrix
for the aggregate communication chart. But feedback during the progress presentation session convinced us that
a chord diagram would suffer from scalability problems. Adding an option for displaying a chord diagram remains
part of future work.

@@

The current implementation focuses on scalability. We managed to some extent to
address the scalability issues raised by big query plans, large amount of
workers or long running queries. For example the \network view uses a
scalable heatmap to show the communication between all the workers in the
system. However, other techniques have to be further implemented to address
this issue. One example would be a more interactive query plan graph, which the
users can zoom. Another example, in the \fragment view that shows spans
for all events we should limit the data that is being downloaded and rendered
to what the user can see.

>>>>

In particular, \system is built to inspect query execution in
Myria\footnote{\url{http://myria.cs.washington.edu/}}, a distributed big data
management system currently being developed in the UW CSE database group. Myria
aims towards building a distributed database platform to provide \emph{big data
management and analytics as a service} primarily for scientific applications.
The proposed visualizations can easily be applied to other DDBMS as well (e.g.
Spark, Hadoop).

@@

Our goal is to build a visualization tool designed to help a database administrator understand the data flow and run-time
performance during query execution in a Distributed Database Management System (DDBMS). Using our tool, the
database programmer or administrator could quickly discover run-time bottlenecks as well as problematic data skew issues,
saving them significant guesswork and programming effort to identify the appropriate logs to analyze. Our visualization tool,
\system, uses profiling logs from \emph{Myria} \cite{myria} as the prototype DDMS. Myria is a distributed
database platform that provides \emph{big data management and analytics as a service} primarily for scientific applications.
Myria takes queries in high-level query languages, such as SQL, translating them into physical query plans and finally executing
them in parallel. Currently, Myria does not offer a visualization tool designed to show the physical query plan, job execution and network
communication between distributed workers. \system allows the programmer/database administrator to visualize the following aspects
of query execution.
\todo{it's not a dag, as we noticed}

@@

Twitter developed Ambrose\cite{ambrose}, a platform for visualization and real-time monitoring of MapReduce data workflows. They offer three different views to show associated jobs, job dependencies and progress. Ambrose has been released as open source. Ambrose, however, is not suitable for our needs as the abstraction level of jobs is too high and does not capture single operators.

Google's Dapper\cite{sigelman2010dapper} is a distributed systems tracing infrastructure and offers fine grained tracing of calls in Google's distributed systems. They also proposed an interface for visualizing traces. Similarly, X-trace\cite{fonseca2007x} was developed as a framework to trace which events cause what other events in a distributed environment. recently, there has been work on visualizing event traces collected in X-trace\footnote{\url{https://github.com/brownsys/X-Trace/tree/master/src/webui/html/interactive}}. Due to the importance of these kinds of debug facilities, Twitter closely modeled Zipkin\cite{zipkin} after Dapper and X-trace and released it as open source.

Dapper and X-trace focus on how data flows through a distributed system. In \system the visualization focuses on the operators and how the data flows through them. This orthogonal view is better suited for debugging performance bottlenecks in DDBMS and also scales to a larger number of events. Furthermore, in contrast to  we offer different abstraction levels, which enables users to find problems faster and handle larger amounts of profiling data. Also, \system is specifically designed to help developers understand query execution in distributed database system like Myria as opposed to general traces in distributed systems.

Tools to visualize query plans for example in SQL Server, as used to improve performance for the SDSS Sky survey\cite{szalay2002sdss}, focus on optimizing queries and not query execution and have no visualization of data flow, which is necessary to optimize physical query execution.

@@

@@

The Worker Communication view consists of two charts: the aggregate communication
chart and the time series chart. The aggregate communication chart is a matrix (Figure~\ref{fig:network}) representing
a cluster of $N$ workers. Each cell $m,n$ in the matrix represents the total number of tuples sent
by worker $m$ to worker $n$ during the query execution. We chose to use the colors from color brewer\cite{harrower2003colorbrewer} and interpolated the colors between the steps using the \emph{CIE L*a*b*} interpolation and lightness correction \cite{chromajsaisch}. The colors work in grayscale and can be interpreted by color blind users.

The matrix chart representing aggregate communication between workers help uncover data skew
issues, as discussed in \S \ref{sec:eval}.

% We initially considered having a chord diagram\footnote{\url{http://bl.ocks.org/mbostock/4062006}} instead of a matrix
%for the aggregate communication chart. But feedback during the progress presentation session convinced us that
%a chord diagram would suffer from scalability problems. Adding an option for displaying a chord diagram remains
%part of future work.

@@

The current implementation focuses on scalability. We managed to some extent to
address the scalability issues raised by big query plans, large amount of
workers or long running queries. For example the \network view uses a
scalable heatmap to show the communication between all the workers in the
system. However, other techniques have to be further implemented to address
this issue. One example would be a more interactive query plan graph, which the
users can zoom. Also, using the ability to query the log data on the backend,
the \fragment view should only download and render what the user can see.
